---
title: "finalproj"
output: html_document
date: '2022-11-25'
---


#https://www.kaggle.com/datasets/kamilpytlak/personal-key-indicators-of-heart-disease

#Personal Key Indicators of Heart Disease
#2020 annual CDC survey data of 400k adults related to their health status





```{r}
library(tidyverse) # for data manipulation
library(tidytext) # for tokenization
library(textstem) # for word stemming and lemmatization
library(caret) # for classification
library(ggplot2)
library(ggbump)
library(Hmisc)
library(caret)
library(caTools)
library(arules)
library(arulesViz)
```


```{r}
setwd ("C:/Users/Hardi/Downloads/" )
data<-read.csv('heart_2020_cleaned.csv')
head(data)
```
#build data set 
```{r}
set.seed(1)
df_1<-subset(data,HeartDisease=='Yes')
df_1<-df_1[sample(nrow(df_1),size=50),]
#random select 50 rows which people have HeartDisease from orginal set

df_2<-subset(data,HeartDisease=='No')
df_2<-df_2[sample(nrow(df_2),size=50),]
##random select 50 rows which people not have HeartDisease from orginal set

df<-rbind(df_1,df_2)
#build data set


```

```{r}
head(df)
```

```{r}
str(df)
```

```{r}
summary(df)
```

```{r}
hist.data.frame(df)
```

#Data pre-processing 
###To deal with BMI 
```{r}
#If your BMI is less than 18.5, it falls within the underweight range. 
#If your BMI is 18.5 to 24.9, it falls within the normal or Healthy Weight range. 
#If your BMI is 25.0 to 29.9, it falls within the overweight range.
#https://www.cdc.gov/healthyweight/assessing/bmi/adult_bmi/
hist(df$BMI)
```

```{r}
df[,'BMI']<-cut(df$BMI,breaks=c(0,18.5,25,30,60),labels = c('under weight','normal weight','over weight','obese weight'))
#rebuild BMI column as define 'underweight','normalweight','overweight','obese weight'
```

```{r}
View(df)
```
###deal with sleep time
```{r}
#Most adults need 7 to 9 hours, although some people may need as few as 6 hours or as many as 10 hours of sleep each day.
#Older adults (ages 65 and older) need 7-8 hours of sleep each day. 
#https://www.cdc.gov/sleep/about_sleep/how_much_sleep.html
hist(df$SleepTime)
```

```{r}
df[,'SleepTime']<-cut(df$SleepTime,breaks=c(0,6,9,12),labels = c('less sleep','good sleep','over sleep'))

```

###deal with PhysicalHealth and MentalHealth
```{r}
hist(df$PhysicalHealth)
```

```{r}
hist(df$MentalHealth)
```

```{r}
df[,'PhysicalHealth']<-cut(df$PhysicalHealth,breaks=c(-1,0,25,30),labels = c('no Physical problem','sometimes Physical problem','always   problem'))
df[,'MentalHealth']<-cut(df$MentalHealth,breaks=c(-1,0,25,30),labels = c('no Mental problem','sometimes Mental problem','always Mental problem'))
```



###Data pre-processing for others
```{r}
df$Smoking<-ifelse(df$Smoking == "Yes", 'smoke', 'not smoke') 
df$AlcoholDrinking<-ifelse(df$AlcoholDrinking == "Yes", 'drink alcohol', 'no drink alcohol') 
df$Stroke<-ifelse(df$Stroke == "Yes", 'have Stroke', 'no Stroke') 
df$DiffWalking<-ifelse(df$DiffWalking == "Yes", 'Diff to Walk', 'Ok to walk') 
df$Diabetic<-ifelse(df$Diabetic == "Yes", 'have Diabetic', 'no Diabetic') 
df$PhysicalActivity<-ifelse(df$PhysicalActivity == "Yes", 'have PhysicalActivity', 'no PhysicalActivity') 
df$Asthma<-ifelse(df$Asthma == "Yes", 'have Asthma', 'no Asthma') 
df$KidneyDisease<-ifelse(df$KidneyDisease == "Yes", 'have KidneyDisease', 'no KidneyDisease') 
df$SkinCancer<-ifelse(df$SkinCancer == "Yes", 'have SkinCancer', 'no SkinCancer') 
#re-name the value in each columns to raplce y/n
```

```{r}
head(df) 
```

```{r}
hist.data.frame(df)
```
#Rules
###RHS HeartDisease
```{r}
rule_rhs = apriori(data = df,
                parameter=list(supp = 0.04,
                               conf = 0.8,
                               minlen = 1),
               appearance = list(default = 'lhs',
                                 rhs = 'HeartDisease=Yes'),
               control = list(verbose=F))
#HeartDisease=Yes as RHS to see what may cause HeartDisease

rule_rhs = sort(rule_rhs,
             by = 'confidence',
             decreasing = TRUE)
#sort by confidence

inspect(rule_rhs[1:20])
```


```{r}
rule_rhs_s<-sort(rule_rhs,
             by = 'support', 
             decreasing = TRUE)
inspect(rule_rhs_s[1:10])
#sort by support
```

```{r}
rule_rhs_l<-sort(rule_rhs,
             by = 'lift', 
             decreasing = TRUE)
inspect(rule_rhs_l[1:10])
#sort by lift
```

### RHS healthy
```{r}
rule_rhs_h = apriori(data = df,
                parameter=list(supp = 0.04,
                               conf = 0.8,
                               minlen = 1),
               appearance = list(default = 'lhs',
                                 rhs = 'HeartDisease=No'),
               control = list(verbose=F))
#HeartDisease=no as RHS to how to be healthy 

rule_rhs_h = sort(rule_rhs_h,
             by = 'confidence',
             decreasing = TRUE)
#sort by confidence
inspect(rule_rhs_h[1:5])
```

###LHS
```{r}
rule = apriori(df,
                parameter = list(supp = 0.05 , conf = 0.8))
rule<-sort(rule,
             by = 'confidence', 
             decreasing = TRUE)
#sort by confidence
inspect(rule[1:20])
```


#graphic
###build new data set
```{r}
df_y<-df[df$HeartDisease=='Yes',]
df_n<-df[df$HeartDisease=='No',]

df_y<-df_y %>%rename(YHeartDisease=HeartDisease,
               YBMI=BMI,
               YSmoking=Smoking,
               YAlcoholDrinking=AlcoholDrinking,
               YStroke=Stroke,
               YPhysicalHealth=PhysicalHealth,
               YMentalHealth=MentalHealth,
               YDiffWalking=DiffWalking,
               YSex=Sex,
               YAgeCategory=AgeCategory,
               YRace=Race,
               YDiabetic=Diabetic,
               YPhysicalActivity=PhysicalActivity,
               YGenHealth =GenHealth ,
               YSleepTime=SleepTime,
               YAsthma=Asthma,
               YKidneyDisease=KidneyDisease,
               YSkinCancer=SkinCancer)

df_n<-df_n %>%rename(NHeartDisease=HeartDisease,
               NBMI=BMI,
               NSmoking=Smoking,
               NAlcoholDrinking=AlcoholDrinking,
               NStroke=Stroke,
               NPhysicalHealth=PhysicalHealth,
               NMentalHealth=MentalHealth,
               NDiffWalking=DiffWalking,
               NSex=Sex,
               NAgeCategory=AgeCategory,
               NRace=Race,
               NDiabetic=Diabetic,
               NPhysicalActivity=PhysicalActivity,
               NGenHealth =GenHealth ,
               NSleepTime=SleepTime,
               NAsthma=Asthma,
               NKidneyDisease=KidneyDisease,
               NSkinCancer=SkinCancer)

df_t<-cbind(df_y,df_n)
head(df_t)
```
### diffs of age 
```{r}
ggplot(data = df, mapping = aes(x=AgeCategory,fill=HeartDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
```
### diffs of age smoking
```{r}
ggplot(data =df_y, mapping = aes(y=YSmoking,fill=YHeartDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
#people have HeartDisease
```

```{r}
ggplot(data =df_n, mapping = aes(y=NSmoking,fill=NHeartDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
#people not have HeartDisease
```
### diffs of BMI
```{r}
ggplot(data = df, mapping = aes(x=BMI,fill=HeartDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
```

### diffs of Alcohol
```{r}
ggplot(data = df, mapping = aes(x=AlcoholDrinking,fill=HeartDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
```

###diffs of 

```{r}
ggplot(data = df, mapping = aes(x=Stroke,fill=HeartDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
```
### diffs of SkinCancer

```{r}
ggplot(data = df, mapping = aes(x=SkinCancer,fill=HeartDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
```

### diffs of MentalHealth

```{r}
ggplot(data = df, mapping = aes(x=MentalHealth,fill=HeartDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
```

### diffs of PhysicalHealth

```{r}

ggplot(data = df, mapping = aes(x=PhysicalHealth,fill=HeartDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
```
### diffs of Diabetic
```{r}
ggplot(data = df, mapping = aes(x=Diabetic,fill=HeartDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
```



##Others graph
###Diff of Alcohol & kidneydisease
```{r}
ggplot(data = df, mapping = aes(x=AlcoholDrinking,fill=KidneyDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
```

```{r}
ggplot(data = df, mapping = aes(x=Smoking,fill=KidneyDisease)) + 
geom_bar(position = 'stack')+
scale_x_discrete(guide=guide_axis(angle=15))
```




#Machine learning
###pre=processing
```{r}
dfr<-df%>%mutate_all(as.factor)
summary(dfr)
```

```{r}
str(dfr)
```

###build test&train set
```{r}
set.seed(666) 
split = sample.split(dfr$HeartDisease, SplitRatio = 0.8) 
  
training = subset(dfr, split == TRUE) 
test = subset(dfr, split == FALSE) 
```
###10-fold cv
```{r}
train.control = trainControl(
  method = 'repeatedcv', 
  number = 10,
  repeats = 10
  )
```

### Random forest
```{r}
search.grid = expand.grid(.mtry = (1:10)) 

rf = train(HeartDisease ~.,
           data = training, 
           method = 'rf',
           metric = 'Accuracy',
           trControl = train.control,
           tuneGrid = search.grid)
```

```{r}
confusionMatrix(rf)
```


```{r}
pred = predict(rf, newdata = test)
confusionMatrix(pred, test$HeartDisease)
```
###SVM
```{r}
search.grid = expand.grid(degree = c(1, 2, 3),scale=c(1),C = seq(0.1, 2, length = 10))

svm = train(HeartDisease ~.,
            data = training,
            method = 'svmPoly',
            trControl = train.control,
            scale = FALSE,
            tuneGrid = search.grid)
```

```{r}
confusionMatrix(svm)
```

```{r}
pred = predict(svm, newdata = test)
confusionMatrix(pred, test$HeartDisease)
```

##Improve
```{r}
dfr<-dfr[,c(-4,-7,-11,-18)]
#delete the seems unrelated columns
```

```{r}
str(dfr)
```
#re-build Ramdom forest
```{r}
set.seed(666) 
split = sample.split(dfr$HeartDisease, SplitRatio = 0.8) 
  
training = subset(dfr, split == TRUE) 
test = subset(dfr, split == FALSE) 
```


```{r}
search.grid = expand.grid(.mtry = (1:10)) 

rf = train(HeartDisease ~.,
           data = training, 
           method = 'rf',
           metric = 'Accuracy',
           trControl = train.control,
           tuneGrid = search.grid)
```

```{r}
confusionMatrix(rf)
```

```{r}
pred = predict(rf, newdata = test)
confusionMatrix(pred, test$HeartDisease)
```
#re-build the SVM
```{r}
search.grid = expand.grid(degree = c(1, 2, 3),scale=c(1),C = seq(0.1, 2, length = 10))

svm = train(HeartDisease ~.,
            data = training,
            method = 'svmPoly',
            trControl = train.control,
            scale = FALSE,
            tuneGrid = search.grid)
```

```{r}
confusionMatrix(svm)
```

```{r}
pred = predict(svm, newdata = test)
confusionMatrix(pred, test$HeartDisease)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```